---
title: "R Coding Exercise"
---

Placeholder file for the future R coding exercise.

```{r}


```

```{r}
library("dplyr")
library("tidyverse")
library("ggplot2")
library("dslabs")
library("renv")
```

```{r}
#look at help file for gapminder data
help(gapminder)
#get an overview of data structure
str(gapminder)
```

```{r}
#get a summary of data
summary(gapminder)
```

```{r}
#determine the type of object gapminder is
class(gapminder)
```

```{r}
africadata = filter(gapminder, continent == 'Africa')
head(africadata)
```

To get overview of africadata

```{r}
str(africadata)
```

Summary of africadata

```{r}
summary(africadata)
```

Infant survival rate

```{r}
infant_survival_rate = africadata %>%
  select(infant_mortality, life_expectancy)
head(infant_survival_rate)
```

Structure of infant survival

```{r}
str(infant_survival_rate)
```

summary of infant survival

```{r}
summary(infant_survival_rate)
```

Infant mortality

```{r}
african_mortality = africadata %>%
  select(population, life_expectancy)
head(african_mortality)
```

African mortality structure

```{r}
str(african_mortality)
```

African mortality summary

```{r}
summary(african_mortality)
```

```{r}
regression = lm(infant_survival_rate$life_expectancy ~ infant_survival_rate$infant_mortality)
rsq <- function(x, y) summary(lm(infant_survival_rate$life_expectancy ~ infant_survival_rate$infant_mortality))$r.squared
r_square = round(rsq(obs, mod), 3)

coeff = round(summary(regression)$coefficients[2],3)
y_int = round(summary(regression)$coefficients[1],3)
plot(
  infant_survival_rate$infant_mortality,
  infant_survival_rate$life_expectancy,
  main='African Life Expectancy \n Infant Mortality Effect',
  xlab='Infant Mortality',
  ylab='Life Expectancy',
   col='red', 
   bg='blue', 
   pch = 21,
)
```

```{r}

```

Life expectancy as a function of population

```{r}
regression = lm(african_mortality$life_expectancy ~ african_mortality$population)
rsq <- function(x, y) summary(lm(african_mortality$life_expectancy ~ african_mortality$population))$r.squared
r_square = round(rsq(obs, mod), 3)

coeff = round(summary(regression)$coefficients[2],9)
y_int = round(summary(regression)$coefficients[1],3)
plot(
  african_mortality$population,
  african_mortality$life_expectancy,
  main='African Life Expectancy \n Population Effect',
  xlab='Population',
  ylab='Life Expectancy',
  col='pink', 
  bg='purple', 
  pch = 21,
)
```

Missing years for infant mortality

```{r}
miss_inf_mort_rows = africadata %>%
  filter(is.na(infant_mortality))
head(miss_inf_mort_rows)
```

Perform year group by to identify missing year

```{r}
miss_inf_mort_yrs = miss_inf_mort_rows %>%
  group_by(year) %>%
  count(infant_mortality)
miss_inf_mort_yrs
```

Extract only year 2000 from data

```{r}
africadata2000 = filter(africadata, year == 2000)
africadata2000
```

Structure of year 2000 data

```{r}
str(africadata2000)
```

Summary of yer 2000 data

```{r}
summary(africadata2000)
```

To create a filter from dataset

```{r}
african_mortality2000 = africadata2000 %>%
  select(population, life_expectancy)
head(african_mortality2000)
```

To create another dataset

```{r}
infant_survival_rate2000 = africadata2000 %>%
  select(infant_mortality, life_expectancy)
head(african_mortality2000)
```

infant mortality effect

```{r}
regression = lm(infant_survival_rate2000$life_expectancy ~ infant_survival_rate2000$infant_mortality)
rsq <- function(x, y) summary(lm(infant_survival_rate2000$life_expectancy ~ infant_survival_rate2000$infant_mortality))$r.squared
r_square = round(rsq(obs, mod), 3)
# summary(regression)$coefficients - keep to sanity check
coeff = round(summary(regression)$coefficients[2],3)
y_int = round(summary(regression)$coefficients[1],3)
plot(
  infant_survival_rate2000$infant_mortality,
  infant_survival_rate2000$life_expectancy,
  main='African Life Expectancy \n Infant Mortality Effect',
  xlab='Infant Mortality',
  ylab='Life Expectancy',
   col='green', 
   bg='violet', 
   pch = 21,
)
```

Population effect

```{r}
regression = lm(african_mortality2000$life_expectancy ~ african_mortality2000$population)
rsq <- function(x, y) summary(lm(african_mortality2000$life_expectancy ~ african_mortality2000$population))$r.squared
r_square = round(rsq(obs, mod), 3)
# summary(regression)$coefficients - keep to sanity check
coeff = round(summary(regression)$coefficients[2],9)
y_int = round(summary(regression)$coefficients[1],3)
plot(
  african_mortality2000$population,
  african_mortality2000$life_expectancy,
  main='African Life Expectancy \n Population Effect',
  xlab='Population',
  ylab='Life Expectancy',
  col='violet', 
  bg='red', 
  pch = 21,
)
```

Simple model fit

Life expectancy to infant mortality

```{r}
fit1 = lm(life_expectancy ~ log(infant_mortality), data=infant_survival_rate2000)
summary(fit1)
```

The R-squared is 0.54, this means that 54% of life expectancy can be explained by infant mortality

Fit life expectancy to population

```{r}
fit2 = lm(life_expectancy ~ log(population), data=african_mortality2000)
summary(fit2)
```

The R-squared us 0.01 which means that 1% of population can be explained by life expentancy
